---
title: "Using Comparator Selection Explorer with Subsets"
author: "James P. Gilbert"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Package Design}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This documents provides a step-by-step guide on using the Comparator Selection Explorer
with subsets in R. It covers setting up the environment, creating cohorts from JSON 
files, and generating cohort tables in a database. The guide also demonstrates how to 
customize cohorts using SQL, RxNorm and ATC templates, and how to leverage the Shiny
explorer for interactive cohort analysis.

#Setup Environment and Load Libraries

```{r}
#Load required libraries
library(Eunomia)
library(DatabaseConnector)
library(CohortGenerator)

#Establish required connection to Eunomia
connectionDetails <- getEunomiaConnectionDetails()
connection <- connect(connectionDetails)
```

#Install OHDSI Cohort Generator (if not already installed)
```{r}
remotes::install_github("OHDSI/CohortGenerator", ref = "develop")
```

#Create Cohort Definitions from JSON files
```{r}
#Initialize empty Cohort Definition set
cohortsToCreate <- CohortGenerator::createEmptyCohortDefinitionSet()

#Load cohort JSON files from package directory
cohortJsonFiles <- list.files(path = system.file("testdata/name/cohorts", 
                                                 package = "CohortGenerator"), full.names = TRUE)

#Loop to read JSONs and create SQL
for (i in 1:length(cohortJsonFiles)) {
  cohortJsonFileName <- cohortJsonFiles[i]
  cohortName <- tools::file_path_sans_ext(basename(cohortJsonFileName))
  # Here we read in the JSON in order to create the SQL
  # using [CirceR](https://ohdsi.github.io/CirceR/)
  # If you have your JSON and SQL stored differenly, you can
  # modify this to read your JSON/SQL files however you require
  cohortJson <- readChar(cohortJsonFileName, file.info(cohortJsonFileName)$size)
  cohortExpression <- CirceR::cohortExpressionFromJson(cohortJson)
  cohortSql <- CirceR::buildCohortQuery(cohortExpression, options= 
                                         CirceR::createGenerateOptions(generateStats = FALSE))
  cohortsToCreate <- rbind(cohortsToCreate, data.frame(cohortId = i,
                                                       cohortName = cohortName, 
                                                       sql = cohortSql,
                                                       stringsAsFactors = FALSE))
}

```

#Create Cohort Tables in Database

```{r}
# Generate the cohort set against Eunomia. 
# cohortsGenerated contains a list of the cohortIds 
# successfully generated against the CDM


# Create the cohort tables to hold the cohort generation results
cohortTableNames <- CohortGenerator::getCohortTableNames(cohortTable = "my_cohort_table")
CohortGenerator::createCohortTables(connectionDetails = connectionDetails,
                                                        cohortDatabaseSchema = "main",
                                                        cohortTableNames = cohortTableNames)
```

#Create Cohorts in Database

```{r}
#Create Empty Cohort Definition Set
cohortDefinitionSet <- createEmptyCohortDefinitionSet()

createCohortTables(connection = connection, cohortDatabaseSchema = "main")
#status <- generateCohortSet(connection = connection,
 #                           cdmDatabaseSchema = "main",
  #                          cohortTableNames = getCohortTableNames(),
   #                         cohortDefinitionSet = cohortDefinitionSet,
    #                        incremental = TRUE,
     #                       incrementalFolder = tempfile())
```

```{r}
# Generate the cohorts
cohortsGenerated <- CohortGenerator::generateCohortSet(
                                          connectionDetails = connectionDetails,
                                          cdmDatabaseSchema = "main",
                                          cohortDatabaseSchema = "main",
                                          cohortTableNames = cohortTableNames,
                                          cohortDefinitionSet = cohortsToCreate)

# View cohort counts
cohortCounts <- CohortGenerator::getCohortCounts(connectionDetails = connectionDetails,
                                                 cohortDatabaseSchema = "main",
                                                 cohortTable = cohortTableNames$cohortTable)
print(cohortCounts)
```




#View Cohort Tables
```{r}
# Define your cohort table name
cohortTable <- getCohortTableNames()$cohortTable

# SQL to select all entries
sql <- paste0(
  "SELECT * FROM main.", cohortTable,
  " WHERE cohort_definition_id = 1"  # assuming cohortId 1
)

# Run the query
cohortEntries <- querySql(connection, sql)

# View the first few entries
head(cohortEntries)
```

```{r}
# Create RxNorm ingredient cohort template
rxNormDefinition <- createRxNormCohortTemplateDefinition(
  connection = connection,  # Testing with Eunomia Database
  cdmDatabaseSchema = "main",
  priorObservationPeriod = 365
)

cohortDefinitionSet <- cohortDefinitionSet |>
        addCohortTemplateDefintion(cohortTemplateDefintion = rxNormDefinition)


# View details of generated template references
rxNormReferences <- rxNormDefinition$getTemplateReferences()
head(rxNormReferences)
```

## ATC Based cohorts
```{r}
# Create ATC-based cohort template
atcDefinition <- createAtcCohortTemplateDefinition(
  connection = connection,  # Replace with your DatabaseConnector connection
  cdmDatabaseSchema = "main",
  mergeIngredientEras = TRUE,
  priorObservationPeriod = 365
)

cohortDefinitionSet <- cohortDefinitionSet |>
        addCohortTemplateDefinition(cohortTemplateDefintion = atcDefinition)

# View ATC template references
atcReferences <- atcDefinition$getTemplateReferences()
head(atcReferences)
```

## Error(s) expected because no ATC in Eunomia dataset


## Running using indication subsets

## Creating Results Database Schema and uploading data

## Using the shiny explorer


