---
title: "Using the Comparator Selection API"
author: "James P. Gilbert"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Using the Comparator Selection API}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteDepends{ComparatorSelectionExplorer}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

The Comparator Selection Explorer ships with a Plumber-based REST API that exposes
cohort search and comparator ranking endpoints. You can host the API on a server
and interact with it programmatically using the built-in R client or any HTTP
client.

## Quick Start

### 1. Host the API

```{r}
library(ComparatorSelectionExplorer)

connectionDetails <- DatabaseConnector::createConnectionDetails(
  dbms = "postgresql",
  server = "localhost/ohdsi",
  user = "user",
  password = Sys.getenv("DB_PASSWORD"),
  cdmSchema = "cdm",
  resultsSchema = "results"
)

startComparatorApi(
  connectionDetails = connectionDetails,
  resultsSchema = "results",
  tablePrefix = "cse_",
  port = 8080
)
```

The server blocks the R session. In production, run it as a background process
or deploy it as a system service (see *Hosting* section below).

### 2. Connect an R Client

From another R session (or a separate script), create a client pointing at the
running server:

```{r}
library(ComparatorSelectionExplorer)

client <- createApiClient(host = "localhost", port = 8080)
```

## API Endpoints

### Health Check

```{r}
client$health()
#> $status
#> [1] "ok"
```

### List Data Sources

```{r}
client$databases()
#>   database_id cdm_source_abbreviation
#> 1   280743270                    CCAE
#> 2   280743271                    MDCR
```

### Search Cohorts

Search all cohorts by name (case-insensitive). Optionally filter by tag
(`"RxNorm"`, `"ATC"`, etc.).

```{r}
# Return all cohorts
all <- client$search()

# Search by keyword
matched <- client$search(q = "ibuprofen")

# Filter by tag
rxnorm_cohorts <- client$search(tag = "RxNorm")

# Both
atc_drugs <- client$search(q = "class", tag = "ATC")
```

### Get Comparator Rankings

This is the primary endpoint. It returns comparators ranked by average cosine
similarity across databases.

```{r}
result <- client$rankings(
  targetCohortId = 101,
  database_ids = c(280743270, 280743271),
  min_databases = 2,
  comparator_type = "RxNorm",
  weight_demo = 20,
  weight_pres = 25,
  weight_hist = 20,
  weight_meds = 20,
  weight_visit = 15
)

result$targetCohortId
#> [1] 101

result$totalComparators
#> [1] 5

result$rankings
#>   rank comparatorId       comparatorName comparatorType avgSimilarity avgCdmRank nDatabases
#> 1    1          102         RxNorm - Drug B        RxNorm         0.874        1.5          2
#> 2    2          202           ATC - Class Y           ATC         0.801        2.5          2
```

#### Parameters

| Parameter | Default | Description |
|-----------|---------|-------------|
| `targetCohortId` | (required) | ID of the target cohort |
| `database_ids` | all | Restrict to specific databases |
| `min_databases` | `2` | Minimum databases a comparator must appear in |
| `comparator_type` | both | `"ATC"`, `"RxNorm"`, or `""` for both |
| `weight_demo` | `20` | Weight for Demographics domain (0-100) |
| `weight_pres` | `20` | Weight for Presentation domain (0-100) |
| `weight_hist` | `20` | Weight for Medical history domain (0-100) |
| `weight_meds` | `20` | Weight for Prior medications domain (0-100) |
| `weight_visit` | `20` | Weight for Visit context domain (0-100) |

Weights are normalised to sum to 1 before scoring.

### Compare Two Cohorts

Retrieve per-domain similarity scores for a specific cohort pair in a specific
database.

```{r}
result <- client$compare(
  cohortId1 = 101,
  cohortId2 = 202,
  database_id = "280743270"
)

result$domainScores
#>        covariateType cosineSimilarity
#> 1       Demographics            0.912
#> 2       Presentation            0.845
#> 3    Medical history            0.768
#> 4         prior meds            0.891
#> 5      visit context            0.722
#> 6            average            0.828
```

## Error Handling

The client throws an error with a descriptive message when the server returns an
HTTP error:

```{r, error = TRUE}
client$search(q = "nonexistent_cohort_xyz")
#> Error: No cohorts match the search criteria
```

Use `tryCatch` in production code:

```{r, error = TRUE}
tryCatch(
  client$rankings(99999, min_databases = 1),
  error = function(e) message("API error: ", e$message)
)
```

## Hosting the API

### Single-session (ad-hoc)

The simplest approach — run `startComparatorApi()` in a terminal or R session:

```r
library(ComparatorSelectionExplorer)

connectionDetails <- DatabaseConnector::createConnectionDetails(
  dbms = "postgresql",
  server = "localhost/ohdsi",
  user = "analyst",
  password = Sys.getenv("DB_PASSWORD"),
  cdmSchema = "cdm",
  resultsSchema = "results"
)

startComparatorApi(
  connectionDetails = connectionDetails,
  resultsSchema = "results",
  tablePrefix = "cse_",
  host = "0.0.0.0",
  port = 8080
)
```

Use `host = "0.0.0.0"` to accept connections from other machines (default is
`"127.0.0.1"` which only allows local connections).

### Background R process (Linux / macOS)

Start the API in a detached R process using `nohup` or `systemd`.

**nohup example:**

```bash
nohup Rscript -e '
  library(ComparatorSelectionExplorer)
  conn <- DatabaseConnector::createConnectionDetails(...)
  startComparatorApi(conn, "results", port = 8080)
' > api.log 2>&1 &
```

### Docker

Build a minimal Docker image:

```dockerfile
FROM rocker/r-ver:4.4
RUN R -e "install.packages(c('ComparatorSelectionExplorer', 'plumber', 'DatabaseConnector'))"
COPY connection_config.rds /opt/api/
CMD R -e "startComparatorApi(readRDS('/opt/api/connection_config.rds'), 'results', port = 8080)"
```

### Process management with `systemd`

Create `/etc/systemd/system/cse-api.service`:

```ini
[Unit]
Description=Comparator Selection Explorer API
After=network.target

[Service]
Type=simple
User=analyst
WorkingDirectory=/opt/cse-api
ExecStart=/usr/bin/Rscript -e '
  library(ComparatorSelectionExplorer)
  conn <- DatabaseConnector::createConnectionDetails(
    dbms = "postgresql",
    server = Sys.getenv("DB_SERVER"),
    user = Sys.getenv("DB_USER"),
    password = Sys.getenv("DB_PASSWORD")
  )
  startComparatorApi(conn, "results", port = 8080)
'
Restart=always
EnvironmentFile=/opt/cse-api/.env

[Install]
WantedBy=multi-user.target
```

Enable and start:

```bash
sudo systemctl daemon-reload
sudo systemctl enable cse-api
sudo systemctl start cse-api
```

### Security considerations

- The API does not include authentication. Protect it with a reverse proxy
  (nginx, Apache) or a firewall.
- Use `host = "127.0.0.1"` (the default) and tunnel through SSH when accessing
  remotely: `ssh -L 8080:localhost:8080 user@server`
- Database credentials are handled by `DatabaseConnector` — keep them out of
  logs and version control.
- The API is read-only: it only queries the results schema and never writes data.
