---
title: "Using Comparator Selection Explorer"
author: "James P. Gilbert, Cameron Atkins"
date: "`r Sys.Date()`"
output:
  html_document:
    number_sections: yes
    toc: yes
  pdf_document:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Package Design}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This document serves as a comprehensive guide to using the Comparator Selection Explorer in R for cohort definition and analysis within the OMOP Common Data Model. It covers setting up the environment, creating cohort templates based on RxNorm ingredients and ATC classifications, generating cohorts in a database, and managing results. Additionally, it demonstrates how to upload results to a database and explore outcomes interactively using a Shiny app. This step-by-step approach enables users to efficiently develop and analyze cohorts for observational research.

# Setup Environment and Load Libraries
This block loads the necessary R packages and establishes a connection to the Eunomia testing database, preparing the environment for cohort creation and exploration.
```{r, message=FALSE}
#Load required libraries
library(Eunomia)
library(DatabaseConnector)
library(CohortGenerator)
library(ComparatorSelectionExplorer)

#Establish required connection to Eunomia
connectionDetails <- getEunomiaConnectionDetails()
connection <- connect(connectionDetails)
```

# Install OHDSI Cohort Generator (if not already installed)
This optional step installs the latest development versions of the OHDSI Cohort Generator and Comparator Selection Explorer packages directly from GitHub.
```{r eval=FALSE}
remotes::install_github("OHDSI/CohortGenerator", ref = "develop")
remotes::install_github("OHDSI/ComparatorSelectionExplorer", ref = "develop")
```

# Create RxNorm Cohort Definitions and view results
This section creates a cohort template based on RxNorm ingredients, adds it to the cohort set, and displays the first few references of the generated template.
```{r, results='hide', message=FALSE}
# Create RxNorm ingredient cohort template
rxNormDefinition <- createRxNormCohortTemplateDefinition(
  connection = connection,  # Testing with Eunomia Database
  cdmDatabaseSchema = "main",
  priorObservationPeriod = 365
)

cohortDefinitionSet <- addCohortTemplateDefintion(cohortTemplateDefintion = rxNormDefinition)


# View details of generated template references
rxNormReferences <- rxNormDefinition$getTemplateReferences()
head(rxNormReferences)
```


# Create ATC Based cohorts
This block creates a cohort template based on ATC drug classifications, adds it to the cohort set, and displays the first few references for review.
```{r eval=FALSE}
# Create ATC-based cohort template
atcDefinition <- createAtcCohortTemplateDefinition(
  connection = connection,  # Replace with your DatabaseConnector connection
  cdmDatabaseSchema = "main",
  mergeIngredientEras = TRUE,
  priorObservationPeriod = 365
)

cohortDefinitionSet <- cohortDefinitionSet |>
        addCohortTemplateDefinition(cohortTemplateDefintion = atcDefinition)

# View ATC template references
atcReferences <- atcDefinition$getTemplateReferences()
head(atcReferences)
```

# Create Cohort Tables in Database
This code generates the necessary cohort tables in the database to store cohort data, using specified schema and table names.
```{r results='hide', message=FALSE}
# Generate the cohort set against Eunomia.
# cohortsGenerated contains a list of the cohortIds
# successfully generated against the CDM


# Create the cohort tables to hold the cohort generation results
cohortTableNames <- CohortGenerator::getCohortTableNames(cohortTable = "my_cohort_table")
CohortGenerator::createCohortTables(connectionDetails = connectionDetails,
                                                        cohortDatabaseSchema = "main",
                                                        cohortTableNames = cohortTableNames)
```

# Generate the Cohort Set and view counts
This section runs the cohort generation process, creating cohorts in the database, and then retrieves and prints the counts of each cohort.
```{r,results='hide', message=FALSE}
# Generate the cohorts
cohortsGenerated <- CohortGenerator::generateCohortSet(
                                          connectionDetails = connectionDetails,
                                          cdmDatabaseSchema = "main",
                                          cohortDatabaseSchema = "main",
                                          cohortTableNames = cohortTableNames,
                                          cohortDefinitionSet = cohortDefinitionSet)

# View cohort counts
cohortCounts <- CohortGenerator::getCohortCounts(connectionDetails = connectionDetails,
                                                 cohortDatabaseSchema = "main",
                                                 cohortTable = cohortTableNames$cohortTable)
print(cohortCounts)
```

# Create Execution Settings
This block sets up the execution environment with all necessary parameters and executes the cohort creation process.
```{r, results='hide', message=FALSE}
executionSettings <- createExecutionSettings(connection = connection,
                                             cohortDefinitionSet = cohortDefinitionSet,
                                             cdmDatabaseSchema = "main",
                                             resultsDatabaseSchema = "main",
                                             cohortTable = cohortTableNames$cohortTable)
execute(executionSettings)
```

# Creating Results Database Schema and uploading data
This code creates a results schema in a SQLite database, prepares it for data, and uploads the cohort results and specifications.
```{r, results='hide', message=FALSE, warning=FALSE}
resultsConnectionDetails <- DatabaseConnector::createConnectionDetails(dbms = "sqlite", server="resultsfile.db")
createResultsDataModel(resultsConnectionDetails, "main", tablePrefix = "")

uploadResults(connectionDetails = resultsConnectionDetails,
              databaseSchema = "main",
              zipFileName = executionSettings$exportZipFile,
              forceOverWriteOfSpecifications = FALSE,
              purgeSiteDataBeforeUploading = FALSE,
              tablePrefix = "")
```


# Using the shiny explorer
This final step launches an interactive Shiny app for exploring the cohort results stored in the database.
```{r eval=FALSE}
launchShinyApp(resultsConnectionDetails, "main")
```



