---
title: "Using Comparator Selection Explorer"
author: "James P. Gilbert"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: yes
  html_document:
    number_sections: yes
    toc: yes
vignette: >
  %\VignetteIndexEntry{Package Design}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
---

This intends to be a brief guide on usage or the comparator selection explorer

ComparatorSelectionExplorer runs three main steps:

1. Generate or load exposure/outcome cohorts.
2. Compute covariate prevalence summaries and cosine similarity scores.
3. Export results and optionally explore them in the Shiny app or through the API.

This vignette shows a reproducible development workflow using Eunomia as the CDM source.

## Create a Eunomia-backed setup

The following example uses Eunomia's synthetic CDM to make local testing and
package evaluation reproducible.

```{r, eval = FALSE}
library(ComparatorSelectionExplorer)
library(Eunomia)
library(CohortGenerator)

# Reproducible local CDM connection.
connectionDetails <- Eunomia::getEunomiaConnectionDetails()

# Example cohort set shipped with CohortGenerator.
cohortDefinitionSet <- CohortGenerator::getCohortDefinitionSet(
  settingsFileName = "Cohorts.csv",
  jsonFolder = "cohorts",
  sqlFolder = "sql/sql_server"
)
```

## Define cohort tag groups

Tag groups are configured with `cohortTags` as a named list of cohort IDs.
These tags constrain cosine similarity calculation.

- Cohorts that share at least one tag are compared.
- Cohorts with no shared tag are not compared.
- If `cohortTags = NULL`, similarity is computed across all eligible cohorts.

```{r, eval = FALSE}
# Example IDs must exist in your cohortDefinitionSet.
# Replace with IDs relevant to your study.
cohortTags <- list(
  nsaid = c(1, 2, 3),
  gi_risk = c(4, 5, 6),
  cardio_risk = c(7, 8)
)
```

## Create execution settings

```{r, eval = FALSE}
# Settings object passed to execute(), exportResults(), and related helpers.
executionSettings <- createExecutionSettings(
  connectionDetails = connectionDetails,
  cdmDatabaseSchema = "main",
  vocabularyDatabaseSchema = "main",
  resultsDatabaseSchema = "main",
  cohortDatabaseSchema = "main",
  cohortTable = "cse_cohort",
  cohortDefinitionSet = cohortDefinitionSet,
  cohortTags = cohortTags,
  minExposureSize = 1000,
  exportZipFile = file.path(tempdir(), "cse_results_main.zip")
)
```

## Execute the workflow

```{r, eval = FALSE}
# Runs cohort generation, feature extraction, cosine similarity, and export.
execute(executionSettings)
```

## Create or migrate results tables and upload

Use these helpers when loading exported zip results into a destination schema.

```{r, eval = FALSE}
# Create or migrate package results tables.
createResultsDataModel(
  connectionDetails = connectionDetails,
  databaseSchema = "main",
  tablePrefix = "cse_"
)

# Upload a previously exported zip file.
uploadResults(
  connectionDetails = connectionDetails,
  databaseSchema = "main",
  zipFileName = executionSettings$exportZipFile,
  tablePrefix = "cse_"
)
```

## Explore results in Shiny

```{r, eval = FALSE}
# Launch UI against populated results tables.
app <- createShinyApp(connectionDetails, "main", tablePrefix = "cse_")
shiny::runApp(app)
```

## Summary

For reproducible local usage:

1. Use Eunomia as the CDM source.
2. Build `executionSettings` with a cohort definition set and optional `cohortTags`.
3. Run `execute()` to produce similarity outputs.
4. Use `createResultsDataModel()` and `uploadResults()` for loading/export workflows.
5. Use `createShinyApp()` (or API endpoints) for interactive exploration.